Designing Modular Software: A Case Study in Introductory Statistics

### 2016-10-10 jcgs-paper-2016/response-1strev.Rmd

---
title: "Response to reviewers"
date: "July 19, 2016"
output: pdf_document
---

@@

(3) The authors ignore other (and at present, much more capable) GUIs for R -- the R Commander and Deducer. Since these are the most direct competitors to intRo -- more so than, e.g., Excel and JMP -- it seems obvious that the authors should address similarities, differences, advantages, and disadvantages. One obvious difference is that these other GUIs run on individual computers rather than being web-based, and it would also make sense to address more generally the advantages and disadvantages of the two approaches. (For example, in my tests of intRo, I noticed slow response, even with very small data sets. Of course, it's hard to know to what to attribute that -- it could as easily have been a network issue as slowness at the server end.)

    **We have included references and a brief discussion on some other GUIs for R, including R Commander and Deducer. We have described some of the differences in our approach to providing a front-end. As for the slow-response time in intRo, this can certainly occur in a couple of different scenarios. One would be if the student is working with a large dataset, as mentioned in Section 5. A second reason for slowdown would be if the student's internet connection is performing slowly. This is an unfortunate limitation of web-based interactive software. However, we have detailed the process by which a student may download and run an intRo instance locally to counteract this possibility.**

# Reviewer 2

(1) The “Article Assessment” questions on the report form don’t work for a paper like this.
The authors report on some very impressive work. The simplicity of conception and execution is one of its biggest strengths. Their motivations are compelling, as is their system design and execution but I think the story-telling can be improved considerably.

@@

    **We appreciate the very kind comments and hope that this revision has improved the story being told.**

(2) I think the authors need to decide, “What sort of paper am I writing?” Included in the manuscript are elements of “a paper about teaching introductory statistics using a software system”, elements of “a paper about designing and building a simple extensible system”, and “a paper about how to use and extend a particular system”. I think the most compelling story for the Journal of Computational and Graphical Statistics is the designing and building of a simple online system for intro stat that is easy to learn, use and extend – and teaches R as a side effect. This would leave room for a sister publication in something like the Journal of Statistics Education which actually focusses on teaching aspects. If you buy this conception, the teaching-related aspects should be shortened, shaved of much detail and form the basis of an introductory motivational section that leads into a design brief and a plan of attack (using shiny etc). There needs to be high-level discussion of how the system is conceived and put together accompanied by high-level diagrams before descending into code fragments (much detailed code-fragment discussion could go to appendices). The paper just needs to convey the sense of how things are done. Actual details of how to customise and extend the system are provided in the system’s documentation.

@@

    **We have taken care to significantly shift the focus away from the teaching aspects, and towards the extensible software system aspects. In particular, we have removed the lengthy teaching discussion and moved elements of it to the motivational section. We have also trimmed much of the content involving supported and unsupported functionality. Specific changes are noted in the responses to the further points raised below.**

(3) Everywhere – distinguish between user=student (who uses what is served up) and user=teacher (who customises what is served up)

@@

    **We went through and replaced instances of "user" with either "student" or "teacher" as appropriate.**

(4) p. 2: Another online analysis system implemented in shiny is iNZight Lite at http://lite.docker.stat.auckland.ac.nz/. It is more ambitious and therefore more complex. Also it does not have the “exposing the R code” capability.

    **We have included a mention of iNZight Lite, along with the shared goals and some differences with intRo.**

(5) p 3 bottom: Remove parentheses around “(and its corresponding code)” because this is actually one of the key features.

@@

    **Done**

(6) R Commander should enter the discussion around here (even though it isn’t online)

@@

    **We have included R Commander in the discussion.**

(7) P 5: Fig 2 shows a simple design that should be easy for casual users (as intro students really are) to get their heads around. Not all that keen on the use of “statistics” as a category heading in Panel 2 because it reinforces the ghettoizing of the word conflicting with a message that says everything you are doing here is statistics.

@@

    **We have updated the application (and the screenshot) to refer to the contingency, regression, and t-test modules category heading as "inference" rather than "statistics", to help emphasize that numerical and graphical summaries are also statistics.**

(8) P 6 top: When you look at the actual R code generated it can be quite complex with its pipe structures etc which I would have thought does rather limit its introductory value.

@@

    **We appreciate the comment. In teaching R Workshops at Iowa State, we observed that among many of the introductory R students we encountered, the pipe structure was often preferred to using traditional nested function calls. The pipe structure is also the standard by which `ggvis` plots are created, which we are using as the backend to our graphical module. We do understand, however, that certain instructors may wish to include a modified version of the default modules removing the piping, which by following the procedure described in the paper would be a relatively simple task.**

(9) P 6-7: discussion of particular modules should be in a “for example” vein. There is no need to justify what particular features are and are not included. (That would be in a teaching paper.)

@@

    **We have removed the in-depth discussion about the particular modules, and instead more vaguely described the functionality. We have also made sure to indicate that the full documentation is available online, which covers the supported functionality of every provided module.**

(10) P 7 top: Being able to choose only plot types that are “compatible” with the variables used does, however, preclude learning experiences of the form “this thing you want to do is not appropriate for this type of data.”

@@

    **We agree. We have restructured this portion of the paper to merely discuss the motivation behind an extensible system, with mentions of disabling and enabling modules. We have moved the details to the Understanding intRo section with the details on modules.**

(12) P 9-10 looks like appendix stuff to me.

    **We have move the implementation details of dynamic UI generation to the appendix.**

(13) Section 3 discusses downloading onto a local machine before customising the hosted online system which should be the focus (again conception and issues involved before any details).

@@

    **We have moved this section to after the discussion on customizing intRo. It now comes just before the conclusion.**

(14) P 11 near bottom: “as an application on the user’s account” puzzled me -- needs to say shinyapps.io account

@@

    **Done**

(15) Section 3.2 mainly belongs in up-front motivation. Table 1 is an implementation mechanism.

    **We have merged remnants of Section 3.2 into Section 2 / motivation. We have also drastically cut down the content to focus this paper more on the extensible systems side, rather than the teaching side. As such, we have also removed Table 1.**

(16) P 14 top: “currently unsupported” justifications are more relevant to a teaching paper than a software design paper.

@@

    **We have removed this comment.**

(17) P 14 Last para Section 3: could go into motivational discussion.

@@

    **We have moved this paragraph to the motivational discussion.**

(18) Section 4.1 is a big part of what I see as the real story but descends to details too quickly.

@@

    **We have eased the transition into this section (now Section 3.1) by talking more generally about the benefits behind a modular structure. We also have made sure to clarify that our concept of modules is not to be confused with Shiny's new modules feature**

(19) Most of Section 4.2 looks to me like more motivation, except the interpolate stuff which is an implementation mechanism

### 2016-10-10 response.Rmd

---
title: "Response to second review"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
---

# Reviewer 1

(1) The paper seems to me inappropriate for the JCGS. Because it focuses purely on software, I think that (once the other issues are addressed) it would be more appropriate for the Journal of Statistical Software or the R Journal.

    **We have refocused the paper on more general computational concepts of reproducibility, reactivity, and modularity, in the context of designing modern statistical software systems. This fits nicely within the aim of JCGS to "extend the use of computational and graphical methods in statistics and data analysis".**

(2) Although the software looks promising, unless I've missed something, it currently has very limited functionality -- indeed, not enough to sustain most introductory statistics courses. This isn't a permanent limitation, however, since the authors have created interesting infrastructure for extending the software. In its current state, the software doesn't quite look ready for prime time.

    **We further highlight in the introduction as well as the beginning of section two that this was our end goal, simplifying software for the user when necessary.**

(3) The authors ignore other (and at present, much more capable) GUIs for R -- the R Commander and Deducer. Since these are the most direct competitors to intRo -- more so than, e.g., Excel and JMP -- it seems obvious that the authors should address similarities, differences, advantages, and disadvantages. One obvious difference is that these other GUIs run on individual computers rather than being web-based, and it would also make sense to address more generally the advantages and disadvantages of the two approaches. (For example, in my tests of intRo, I noticed slow response, even with very small data sets. Of course, it's hard to know to what to attribute that -- it could as easily have been a network issue as slowness at the server end.)

    **We have added comparisons to more software, including Radiant, to the introduction. Within these comparisons, we focused on extensibility, modularity, reproducibility, licensing, and the functionality of each piece of software. We have also included a table that summarises these comparisons.**

# Reviewer 2

(1) While the authors have made a pretty good fist of addressing everything I asked for and it is pretty
well good to go from my point of view. But I think they can still strengthen their case and their appeal somewhat. I think a title like the following would suggest better the new focus of the paper:
"Customisable introductory statistics software via intRo"

    **We agree, thanks for the suggestion. We have changed the title to "Designing Modular Software: A Case Study in Introductory Statistics", which we feel highlights the customizable aspects of intRo.**

(3) Also there are 2 big ideas in here about software that go beyond the particular application:
extensibility (the ability to “easily” add functionality) and customisation/simplification (the ability to
easily hide functionality) and a way of implementing these. I think these 2 elements and their
advantages should be highlighted from the get go, starting with the abstract. The advantages of
extensibility are obvious but the desirability of being able to simplify the view of software a
particular audience gets needs highlighting. (Also the idea that they enable this by the way in which
the instance of the software that will be served up to the user is assembled on the fly at run time.)
Detailed comments follow. (Perhaps unfortunately, many draw on what I have said above.):

    **We have updated the abstract to reflect the new focus of the paper, which is writing extensible software through the use of modular programming within the framework of shiny. Additionally, we've highlighted the ability to remove functionality and present a simplified view to the user in the abstract. The paper itself now addressed the "big ideas" of extensibility and customization much more thoroughly.**

(4) Page 2 bottom: could note that having a large group of students install software on their own
machines can be a nightmare with differing hardware configurations, versioning etc.

    **We added a sentence in section 2, page 5 top to address this.**

(5) Page 3, last para pf Section 1 needs breaking up.

    **The content of this paragraph is no longer in the paper as a single paragraph, but rather redispersed into multiple sections.**

(6) Page 4, Note that Fig 1 is a schematic, not a view of the software.

    **We have added the verbiage "schematic" in the text and figure caption to note this.**

(7) Page 7 last line: no need for “introductory” here. These ideas apply generally.

    **Good point. We removed the word "introductory" from the sentence.**

(8) Page 8 ff (also page 10-11): I think this story would be better spun in terms of someone writing an
incorporating a new module (a few tiny wording changes).

    **We changed the wording to indicate the example walks through the process of creating a module, as well as including it.**

# Associate Editor

(1) The authors have responded to some of the reviewers' comments, but there is still more work to do. The motivation for intRo is as a simpler interface to R (for teaching statistics AND generating interest in R), but this should just be background information for this article.  There are many other R GUIs that are MUCH more developed for teaching statistics (that also generate R code as well).  The features of intRo that are for the student/user are less interesting and do not add anything beyond existing software. The main focus of the article should be on the novel aspects of this work, which is the modularity/extensibility - the features of intRo that are for the teacher/developer.

    **We have refocused the article onto the novel aspects of modularity and extensibility by reqriting the introduction and reframing the article as a case study in modularity. Accordingly, we have moved details about intRo as an interface to R to section 2.**

(2) The comparison with existing software (Rcmdr, Deducer, iNZight, Shiny) should be more in-depth, but focused on the modularity/extensibility issue.

    **We have added comparisons to more software, including Radiant, to the introduction. Within these comparisons, we focused on extensibility, modularity, reproducibility, licensing, and the functionality of each piece of software. We have also included a table that summarises these comparisons.**

### 2016-10-10 jcgs-paper-2016/paper.Rmd

---
title: "Designing Modular Software: A Case Study in Introductory Statistics"
author:
  - name: Eric Hare
    affiliation: Iowa State University
    department: Department of Statistics
  - name: Andee Kaplan
    affiliation: Iowa State University
    department: Department of Statistics
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: preamble.tex
    number_sections: yes
    template: jcgs_article.tex
fontsize: 12pt
abstract: |
   Modular programming is a development paradigm that emphasizes self-contained, flexible, and independent pieces of functionality. However, the rise of web-based software applications has presented new challenges for designing an extensible, modular software system. In this paper, we outline a framework for designing such a system, with a focus on reproducibility of the results. We present as a case study a Shiny-based web application called `intRo`, that allows the user to perform basic data analyses and statistical routines. Finally, we highlight some challenges we encountered when combining modular programming concepts with reactive programming as used by Shiny.
keywords: Interactivity, Modularity, Programming Paradigms, Reactive Programming, Reproducibility, Statistical Software
bibliography: refs.bib
blind: false
---

>>>>

---
title: "Designing Modular Software: A Case Study in Introductory Statistics"
author:
  - name: Eric Hare
    affiliation: Iowa State University
    department: Department of Statistics
  - name: Andee Kaplan
    affiliation: Iowa State University
    department: Department of Statistics
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: preamble.tex
    number_sections: yes
    template: jcgs_article.tex
fontsize: 12pt
abstract: |
   Modular programming is a development paradigm that emphasizes self-contained, flexible, and independent pieces of functionality. This practice allows new functionality to be seamlessly added when new functionality is desired, and unwanted functionality to be removed, thus simplify the user-facing view of the software. However, the rise of web-based software applications has presented new challenges for designing an extensible, modular software system. In this paper, we outline a framework for designing such a system, with a focus on reproducibility of the results. We present as a case study a Shiny-based web application called `intRo`, that allows the user to perform basic data analyses and statistical routines. Finally, we highlight some challenges we encountered, and how to address them, when combining modular programming concepts with reactive programming as used by Shiny.
keywords: Interactivity, Modularity, Programming Paradigms, Reactive Programming, Reproducibility, Statistical Software
bibliography: refs.bib
blind: false
---
